MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-04-12 latex/paper/paper.tex
o

    \begin{figure}[]
        \centering
        \includegraphics[width=9.0cm]{figures/hierarchy.png}
        \caption{Architecture Hierarchy of a Xilinx FPGA (Ultrascale)}
        \label{fig:hierarchy}
    \end{figure}

@@

    % Figure~\ref{fig:hierarchy} illustrates the architectural hierarchy of an Ultrascale Xilinx FPGA. 
    % The architecture hierarchy is the same for the 7-Series devices. 
    Figure \ref{fig:hierarchy_3} illustrates the architectural hierarchy of a 7-Series Xilinx FPGA. 
    At its most basic level, an FPGA is a vast physical array of replicated atomic components known as Basic Elements of Logic (BELs). 
    These BELs, which include LUTs, FFs, BRAMs, DSPs, and programmable interconnects, form the fundamental building blocks for custom digital circuit implementations on FPGAs.

@@

    \begin{figure}[]
        \centering
        \includegraphics[width=9.0cm]{figures/hierarchy_3.png}
        \caption{Tiles, Sites, BELs hierarchy}
        \label{fig:hierarchy_3}
    \end{figure}

@@

    \begin{figure}
        \centering
        \includegraphics[width=7.0cm]{figures/design_flow.png}
        \caption{A typical FPGA design and verification workflow}
        \label{fig:design_flow}
    \end{figure}

    Figure \ref{fig:design_flow} shows a typical FPGA design and verification workflow.
    The FPGA design flow takes a high-level digital design written in an HDL and generates a configuration file that programs the physical FPGA.

    The process begins with \textbf{design entry}, where an engineer describes the functionality of their circuit using a hardware description language (HDL) such as Verilog or VHDL. This description abstracts away details of the physical hardware while defining the logic and behavior of the design.

    Next, the \textbf{synthesis} tool interprets the HDL code and generates a netlist that represents the design in terms of primitive cells, such as lookup tables (LUTs) and flip flops. 
    This is the result of the synthesis tool parsing the raw Verilog source files, inferring logical behavior, and instantiating primitive cells to emulate the desired logical behavior. 
    The final stage is often referred to as "technology mapping", as the primitive cells are specific to the FPGA architecture and consequently the FPGA manufacturer.

@@

    Placement and routing are generally referred to together as the \textbf{implementation} stage, as both stages are interdependent and must adhere to the constraints of the same device. 
    State of the art (SOTA) tools will perform routing-aware placement for greater optimization.

@@

>>>>

    % \begin{figure}[]
    %     \centering
    %     \includegraphics[width=9.0cm]{figures/hierarchy.png}
    %     \caption{Architecture Hierarchy of a Xilinx FPGA (Ultrascale)}
    %     \label{fig:hierarchy}
    % \end{figure}

    \begin{figure*}[]
        \centering
        \includegraphics[width=15.0cm]{figures/hierarchy_3.png}
        \caption{Architecture Hierarchy of a Xilinx FPGA}
        \label{fig:hierarchy_4}
    \end{figure*}

@@

    % Figure~\ref{fig:hierarchy} illustrates the architectural hierarchy of an Ultrascale Xilinx FPGA. 
    % The architecture hierarchy is the same for the 7-Series devices. 
    Figure \ref{fig:hierarchy_4} illustrates the architectural hierarchy of a 7-Series Xilinx FPGA. 
    At its most basic level, an FPGA is a vast physical array of replicated atomic components known as Basic Elements of Logic (BELs). 
    These BELs, which include LUTs, FFs, BRAMs, DSPs, and programmable interconnects, form the fundamental building blocks for custom digital circuit implementations on FPGAs.

@@

    Note that in the Xilinx architectures, "CLBs" refer to CLB Tiles, each containing two "SLICE" Sites, with each SLICE Site containing 8 LUTs, 8 Flip Flop, and 1 CARRY BEL.
    These SLICEs facilitate the core reprogrammability of the FPGA.

@@

    \begin{figure}
        \centering
        \includegraphics[width=8.0cm]{figures/design_flow.png}
        \caption{A typical FPGA design and verification workflow}
        \label{fig:design_flow}
    \end{figure}

    Figure \ref{fig:design_flow} illustrates a typical FPGA design and verification workflow, transforming a high-level digital design into a configuration file that programs the physical FPGA.

    The process begins with \textbf{design entry}, where an engineer describes the desired functionality using a hardware description language (HDL) such as Verilog or VHDL. 
    Engineers may use various coding styles - behavioral, structural, dataflow, or a mix of all three - to capture the logic and behavior of the design while abstracting away the intricacies of the physical hardware architecture.

    Next, the user runs the \textbf{synthesis} tool, which parses the raw HDL source files to infer logical behavior, perform optimizations, and instantiate primitive cells. 
    The result is a detailed netlist, typically output as a low-level Verilog or Electronic Design Interchange Format (EDIF) file, where the design is represented purely in structural HDL. 
    The final stage is often referred to as "technology mapping", as the vocabulary or library of primitive cells must be specific to the FPGA architecture and consequently the FPGA manufacturer.

@@

    Placement and routing are generally referred jointly as the \textbf{implementation} stage, as both stages are interdependent and must adhere to the constraints of the same device. 
    State of the art (SOTA) tools will perform routing-aware placement for greater optimization.

@@

        \begin{figure}
            \centering
            \includegraphics[width=4.0cm]{figures/edif_net.png}
            \includegraphics[width=4.0cm]{figures/intrasite_route.png}
            \includegraphics[width=8.0cm]{figures/intersite_route.png}
            \caption{
                \textbf{Upper Left}: A LUT-FF cell pair in the netlist.
                \textbf{Upper Right}: The LUT-FF pair placed within the same Site.
                \textbf{Bottom}: The LUT-FF pair placed across different Sites.
            }
            \label{fig:intrasite_intersite}
        \end{figure}

        For example, in many HDL designs, LUTs and FFs often come in pairs, as is the case in Figure \ref{fig:intrasite_intersite}
